---
title: "unique_figures_18S"
author: "Beth"
date: "6/7/2021"
output: html_document
---

Read in and combine unique files for eukarya

```{r}
library(vegan)
library(phyloseq)
library(ggplot2)
#library(DESeq2)
library(tidyverse)
library(rstatix)
library(ggpubr)
library(RColorBrewer)
library(plyr)
library(pheatmap)
library(cowplot)
library(reshape2)

#bacteria
unique <- read.csv('2021.02.15_dawson.eukarya.unique_tally.csv', header = T, row.names = 1)

## convert all na's to 0
unique[is.na(unique)] <- 0
```

Read in and combine taxa & map files  also sample metadata. 

```{r}

taxa <- read.csv('2021.02.15_dawson.eukarya.taxon_map.csv', header = T, row.names = 1, as.is = T)
map <- read.csv('2021.02.15_dawson.eukarya.seq_edge_map.csv',header = T, row.names = 1, as.is = T)


#sample metadata
metadata <- read.csv('dawson_metadata.csv',header = T)
metadata <- metadata[1:63 ,]

#create mapfile combining taxa and map
mapfile <- map[colnames(unique),]
Yep <- map[colnames(unique), 'global_edge_num']
tax <- taxa[Yep, 'taxon']
tax[tax == ""] <- 'Eukarya'
tax[is.na(tax)] <- 'Eukarya'
mapfile$Taxa <- tax
Fam <-taxa[Yep, 'family']
Fam[Fam == ""] <- 'Eukarya'
Fam[is.na(Fam)] <- 'Eukarya'
mapfile$Family <- Fam
Phylum <- taxa[Yep, 'division']
Phylum[Phylum == ""] <- 'Eukarya'
Phylum[is.na(Phylum)] <- 'Eukarya'
mapfile$Phylum <- Phylum
rm(Fam, Phylum, Yep)

```

unique.qc - change sample names from paprica files, change ASV names to unique IDs and remove Cyanos (as they are plastids, not bacteria)

```{r}
#remove cyanos
unique.qc <- unique

#rename row names
rownames(metadata) <- metadata$paprica_file_18S
rownames(unique.qc) <- metadata[rownames(unique.qc), 'SampleID']
rownames(metadata) <- metadata$SampleID

               
```


unique.select - Limit to samples of interest, remove bad library builds, change ASVs to sequence names

```{r}
#limit to station B and carboy samples
unique.select <- unique.qc[grep('B|Carboy|ice|core|Hero|slush|CTD', rownames(unique.qc)),]


## remove any low abundance samples (i.e. bad library builds), and also low abundance reads

unique.select <- unique.select[rowSums(unique.select) > 1,]
unique.select <- unique.select[,colSums(unique.select) > 0]

#normalize by sample abundace

unique.s <- unique.select

unique.select <- unique.select/rowSums(unique.select)

unique.s2 <- unique.s %>%
   mutate(SampleID = rownames(unique.s))

write.csv(unique.s2, file="18S_unique_raw_cleaned.csv")


```


Phylum level bar plot


```{r}
library(ggplot2)
library(tidyr)
#change ASV sequences to Phylum classification 
s <- unique.select
index <- as.data.frame(colnames(s))
index$phylum <- mapfile[index$`colnames(s)`, 'Phylum']
colnames(s) <- 1:length(s)

#melt for bar plot

s.melt <- melt(s)
s.melt$SampleID <- rownames(s)
s.melt <- s.melt %>% filter(value != 0)
s.melt$taxa <- index[s.melt$variable, 'phylum']


bp <- function(x, y){
ggplot() +
geom_bar(aes( y = value, x = SampleID, fill= taxa), data = x, stat = "identity")+ 
  theme(axis.text.x = element_text(angle = 90)) +
   labs(title = y) +
   labs(x = "", y = "Relative Abundance (%)", fill = "class") 
}  

bp(s.melt, "18S Relative Abundance")

```
top 20 taxa heatmap

```{r}
r <- unique.select
colnames(r) <- mapfile[colnames(r), 'Taxa']

hm <- function(x,y) {
x <- x[, colSums(x!=0) > 0]
top_20 <- names(sort(colSums(x), decreasing =TRUE)[1:20])
tally.top <- x[top_20]
pheatmap(t(tally.top), main = y, cex = 1)
}

hm(r, "16S Most Abundant Taxa")
```

diversity indexes with anova

```{r}
library(vegan)
library(ggplot2)

#restrict metadata to useful samples 

data <- metadata[grep('B|Carboy', rownames(metadata)),]
data[,1:3] <- NULL
data <- data[-1,] 

#shannon diversity index
shann <- diversity(unique.s) #use data from BEFORE normalizing

# Simpson index
simp <- diversity(unique.s, "simpson")  

#inverse Simpson
insimp <- diversity(unique.s, "invsimpson")

par(mfrow = c(1, 3))  # to generate panels with 1 row of 3 graphs
hist(shann)
hist(simp)
hist(insimp)

#add inverse simpson diversity metric (looked best) to metadata
a <- as.data.frame(insimp)
data$InvSimpson  <- a[rownames(data), 'insimp']


#plots divided by different sampling metrics (not sure which you think is most important)
ggplot(data, aes(x = as.factor(TempC), y = InvSimpson, fill=Sample_type)) + 
  geom_boxplot()

ggplot(data, aes(x = as.factor(Salinity), y = InvSimpson, fill=Sample_type)) + geom_boxplot()

anova_result <- aov(InvSimpson ~ as.factor(TempC), data)
summary(anova_result)

#not significant (barely), could run a Tukey (the negative temp looks like it could be significant)

```

CA

```{r}
library(ca)
library(factoextra)
#hellinger transform data
Hellinger_unique <- decostand(unique.s, method = "hellinger", na.rm = TRUE) 
Hellinger_unique <- Hellinger_unique[ order(row.names(Hellinger_unique)), ]
#ca
unique.select.ca <- ca(Hellinger_unique)
unique.select.ca
unique.select.ca.var <- get_eigenvalue(unique.select.ca)
unique.select.ca.res <- get_ca_col(unique.select.ca)
samples.x <- unique.select.ca$rowcoord[,1]
samples.y <- unique.select.ca$rowcoord[,2]

dim.1.var <- round(unique.select.ca.var$variance.percent[1], 1)
dim.2.var <- round(unique.select.ca.var$variance.percent[2], 2)
envi_color<-c(seawater='purple', low='orange', mid='green', high='pink')
plot(samples.x, samples.y,
ylab = paste0('Dim 2: ', dim.2.var, '%'),
xlab = paste0('Dim 1: ', dim.1.var, '%'),
pch = 21,col="black", cex=1, bg=envi_color[data$location])
row$coord

#remove outliers


```






NMDS 

```{r}
library(phyloseq)
####################### Ordination Plots ############################
Hellinger_unique <- decostand(unique.s, method = "hellinger", na.rm = TRUE) 

#transform dataframes into matrices, switching the rows and columns for downstream analyses
Hellinger_unique <- t(Hellinger_unique) #flip

#create OTU matrix
OTUmat <- as.matrix(Hellinger_unique)
#create taxa matrix
taxmat <- as.matrix(mapfile[,4])

#combine into a phyloseq object
OTU = otu_table(OTUmat, taxa_are_rows = TRUE)
TAX = taxa_names(taxmat)
sampledata <- data
physeq = phyloseq(OTU,TAX,sampledata)
physeq

#ordination 
theme_set(theme_bw())
ord2 <- ordinate(physeq, "NMDS", "bray")

plot_ordination(physeq, ord2, type="sample_type", color = "TempC") +       geom_point(size=3) + 
   facet_wrap(~TempC, labeller = depth_labeller) +            theme(legend.position = "top", legend.title = element_blank()) + scale_shape_discrete(labels = c("Shaved", "Not Shaved"))


#permutation MANOVA vegan package
df = data.frame(sample_data(physeq))
d = phyloseq::distance(physeq, "bray")

adonis(formula = d ~ treatment*shave + depth, data = df) 
#Homogeneity of dispersion test
beta <- betadisper(d, df$treatment)
permutest(beta) #betadisper results are not sig, meaning we reject null hypothesis that tour groups have the same dispersions... confident in result

#p.adj to correct for multiple comparisons
p_perm <- c(0.001, 0.0152, 0.001, 0.133)
p.adjust(p_perm, "bonferroni")


```
