---
title: "unique_figures_18S"
author: "Beth"
date: "6/7/2021"
output: html_document
---

Read in and combine unique files for eukarya

```{r}
library(vegan)
library(phyloseq)
library(ggplot2)
#library(DESeq2)
library(tidyverse)
library(rstatix)
library(ggpubr)
library(RColorBrewer)
library(plyr)
library(pheatmap)
library(cowplot)
library(MASS)
library(reshape2)
library(reshape)
library(dplyr)

#bacteria
unique <- read.csv('2021.02.15_dawson.eukarya.unique_tally.csv', header = T, row.names = 1)

## convert all na's to 0
unique[is.na(unique)] <- 0
```

Read in and combine taxa & map files  also sample metadata. 

```{r}

taxa <- read.csv('2021.02.15_dawson.eukarya.taxon_map.csv', header = T, row.names = 1, as.is = T)
map <- read.csv('2021.02.15_dawson.eukarya.seq_edge_map.csv',header = T, row.names = 1, as.is = T)


#sample metadata
metadata <- read.csv('dawson_metadata.csv',header = T)
metadata <- metadata[1:63 ,]

#create mapfile combining taxa and map
mapfile <- map[colnames(unique),]
Yep <- map[colnames(unique), 'global_edge_num']
tax <- taxa[Yep, 'taxon']
tax[tax == ""] <- 'Unclassified Eukaryota'
tax[is.na(tax)] <- 'Unclassified Eukaryota'
mapfile$Taxa <- tax
Fam <-taxa[Yep, 'family']
Fam[Fam == ""] <- 'Unclassified Eukaryota'
Fam[is.na(Fam)] <- 'Unclassified Eukaryota'
mapfile$Family <- Fam
Phylum <- taxa[Yep, 'division']
Phylum[Phylum == ""] <- 'Unclassified Eukaryota'
Phylum[is.na(Phylum)] <- 'Unclassified Eukaryota'
mapfile$Phylum <- Phylum

#if problem arises from mapfile, comment out from here below
Kingdom <- taxa[Yep, 'kingdom']
Kingdom[Kingdom == ""] <- 'Unclassified Eukaryota'
Kingdom[is.na(Kingdom)] <- 'Unclassified Eukaryota'
mapfile$Kingdom <- Kingdom

Supergroup <- taxa[Yep, 'supergroup']
Supergroup[Supergroup == ""] <- 'Unclassified Eukaryota'
Supergroup[is.na(Supergroup)] <- 'Unclassified Eukaryota'
mapfile$Supergroup <- Supergroup

Class <- taxa[Yep, 'class']
Class[Class == ""] <- 'Unclassified Eukaryota'
Class[is.na(Class)] <- 'Unclassified Eukaryota'
mapfile$Class <- Class

Order <- taxa[Yep, 'order']
Order[Order == ""] <- 'Unclassified Eukaryota'
Order[is.na(Order)] <- 'Unclassified Eukaryota'
mapfile$Order <- Order

Genus <- taxa[Yep, 'genus']
Genus[Genus == ""] <- 'Unclassified Eukaryota'
Genus[is.na(Genus)] <- 'Unclassified Eukaryota'
mapfile$Genus <- Genus

Species <- taxa[Yep, 'species']
Species[Species == ""] <- 'Unclassified Eukaryota'
Species[is.na(Species)] <- 'Unclassified Eukaryota'
mapfile$Species <- Species

rm(Fam, Phylum, Yep, Kingdom, Supergroup, Class, Order, Genus, Species)

```

unique.qc - change sample names from paprica files, change ASV names to unique IDs and remove Cyanos (as they are plastids, not bacteria)

```{r}
#remove cyanos
unique.qc <- unique

#rename row names
rownames(metadata) <- metadata$paprica_file_18S
rownames(unique.qc) <- metadata[rownames(unique.qc), 'Better_SampID']
rownames(metadata) <- metadata$Better_SampID

               
```


unique.select - Limit to samples of interest, remove bad library builds, change ASVs to sequence names

```{r}
#limit to station B and carboy samples
unique.select <- unique.qc[grep('B|ppt|Ev|Hero', rownames(unique.qc)),]
unique.select <-  unique.select %>% mutate(SampleID = rownames(unique.select))%>%
       filter(!str_detect(SampleID, "EvXSW_A|Ev51Slush_A|Ev15SW_A|B1_1" ))%>%
       dplyr::select(-SampleID)


## remove any low abundance samples (i.e. bad library builds), and also low abundance reads, reorder alphabetically

unique.select <- unique.select[rowSums(unique.select) > 1,]
unique.select <- unique.select[,colSums(unique.select) > 0]
unique.select <- unique.select[order(rownames(unique.select)),]

#normalize by sample abundace

unique.s <- unique.select

unique.select <- unique.select/rowSums(unique.select)

#calculate total reads
q <- unique.s
colnames(q) <- mapfile[colnames(q), 'Taxa']

q.fix <-as.data.frame(t(rowsum(t(q), group = colnames(q), na.rm = T)))
q.fix <- q.fix %>%
  mutate(Total = rowSums(q.fix))

#this gives sample read totals
q.fix$Total

#this gives read totals across all samples
Total_reads <- colSums(q.fix)
Total_reads

#Write out wide data with cleaned ASVs, raw abundance
wide_abu <- unique.s %>%
  mutate(SampleID = rownames(unique.s))

write.csv(wide_abu, file="18S_unique_raw_cleaned.csv")

#write out clean data ASVs with all taxonomic assignments, relative abundance
abu_ID <- t(unique.select)
abu_ID_join <- merge(abu_ID, mapfile, by="row.names")

write.csv(abu_ID_join, file="18S_unique_rel_ID.csv")



```


Phylum level bar plot


```{r}
library(ggplot2)
library(tidyr)
#change ASV sequences to Phylum classification 
s <- unique.select
index <- as.data.frame(colnames(s))
index$phylum <- mapfile[index$`colnames(s)`, 'Phylum']
colnames(s) <- 1:length(s)

#melt for bar plot

s.melt <- melt(s)
s.melt$SampleID <- rownames(s)
s.melt <- s.melt %>% filter(value != 0)
s.melt$taxa <- index[s.melt$variable, 'phylum']
subjmeans <- cast(s.melt, SampleID~taxa, sum)

#write out csv of wide phyla data for CCA
write.csv(subjmeans, file="18S_phyla.csv")

#make long for bar plot
sum.long <- melt(subjmeans, id.vars=c("SampleID", "taxa"))

bp <- function(x, y){
ggplot() +
geom_bar(aes( y = value, x = SampleID, fill= taxa), data = x, stat = "identity")+ 
  theme(axis.text.x = element_text(angle = 90)) +
   labs(title = y) +
   labs(x = "", y = "Relative Abundance (%)", fill = "class") 
}  

bp(sum.long, "18S Relative Abundance")

bplot <- bp(sum.long, "18S Relative Abundance")
save_plot("Euk_barplot.pdf", bplot, base_height = 10, base_width = 10, units="in")

#use similar code to get top supergroups
s <- unique.select
index <- as.data.frame(colnames(s))
index$phylum <- mapfile[index$`colnames(s)`, 'Supergroup']
colnames(s) <- 1:length(s)


#melt for bar plot

s.melt <- melt(s)
s.melt$SampleID <- rownames(s)
s.melt <- s.melt %>% filter(value != 0)
s.melt$taxa <- index[s.melt$variable, 'phylum']
subjmeans <- cast(s.melt, SampleID~taxa, sum)

write.csv(subjmeans, file="18S_supergroup.csv")


#make long for bar plot
sum.long <- melt(subjmeans, id.vars=c("SampleID", "taxa"))

bp <- function(x, y){
ggplot() +
geom_bar(aes( y = value, x = SampleID, fill= taxa), data = x, stat = "identity")+ 
  theme(axis.text.x = element_text(angle = 90)) +
   labs(title = y) +
   labs(x = "", y = "Relative Abundance (%)", fill = "class") 
}  

bp(sum.long, "18S Relative Abundance")

bplot <- bp(sum.long, "18S Relative Abundance")
save_plot("Euk_barplot.pdf", bplot, base_height = 10, base_width = 10, units="in")


#use similar code to get top class
s <- unique.select
index <- as.data.frame(colnames(s))
index$phylum <- mapfile[index$`colnames(s)`, 'Class']
colnames(s) <- 1:length(s)

#melt for bar plot

s.melt <- melt(s)
s.melt$SampleID <- rownames(s)
s.melt <- s.melt %>% filter(value != 0)
s.melt$taxa <- index[s.melt$variable, 'phylum']
subjmeans <- cast(s.melt, SampleID~taxa, sum)

write.csv(subjmeans, file="18S_class.csv")


#use similar code to get top genus
s <- unique.select
index <- as.data.frame(colnames(s))
index$phylum <- mapfile[index$`colnames(s)`, 'Genus']
colnames(s) <- 1:length(s)

#melt for bar plot

s.melt <- melt(s)
s.melt$SampleID <- rownames(s)
s.melt <- s.melt %>% filter(value != 0)
s.melt$taxa <- index[s.melt$variable, 'phylum']
subjmeans <- cast(s.melt, SampleID~taxa, sum)

write.csv(subjmeans, file="18S_genus.csv")


##try making barplot of diatoms only
abu_diatom <- abu_ID_join %>%
  filter(str_detect(Class, "Bacillariophyta"))

#take out extra id info for now
abu_diatom <- abu_diatom[,1:33]

#make row names back into row names
rownames(abu_diatom) <- abu_diatom[,1]
abu_diatom[,1] <- NULL

#use similar code to get top genus diatoms
s <- as.data.frame(t(abu_diatom))
index <- as.data.frame(colnames(s))
index$phylum <- mapfile[index$`colnames(s)`, 'Genus']
colnames(s) <- 1:length(s)

#melt for bar plot

s.melt <- melt(s)
s.melt$SampleID <- rownames(s)
s.melt <- s.melt %>% filter(value != 0)
s.melt$taxa <- index[s.melt$variable, 'phylum']
subjmeans <- cast(s.melt, SampleID~taxa, sum)

write.csv(subjmeans, file="18S_diatom_genus.csv")

#make version filled so all out of 1 (total diatom proportion)
subjmeans_prop <- data.stand((as.matrix(subjmeans)), method='total', margin='row', plot=F)
write.csv(subjmeans_prop, file="18S_diatom_genus_proportional.csv")

#make long for bar plot
sum.long <- melt(subjmeans, id.vars=c("SampleID", "taxa"))

bp <- function(x, y){
ggplot() +
geom_bar(aes( y = value, x = SampleID, fill= taxa), position = "fill", data = x, stat = "identity")+ 
  theme(axis.text.x = element_text(angle = 90)) +
   labs(title = y) +
   labs(x = "", y = "Relative Abundance (%)", fill = "class") 
}  

bp(sum.long, "18S Relative Abundance")

bplot <- bp(sum.long, "18S Relative Abundance")
save_plot("Euk_barplot_diatoms.pdf", bplot, base_height = 10, base_width = 10, units="in")

#make version filled so all out of 1 (total diatom proportion)
subjmeans_prop <- data.stand((as.matrix(subjmeans)), method='total', margin='row', plot=F)
write.csv(subjmeans_prop, file="18S_diatom_genus_proportional.csv")


##try making barplot of dinos only
abu_dino <- abu_ID_join %>%
  filter(str_detect(Class, "Dinophyceae"))

#take out extra id info for now
abu_dino <- abu_dino[,1:33]

#make row names back into row names
rownames(abu_dino) <- abu_dino[,1]
abu_dino[,1] <- NULL

#use similar code to get top genus diatoms
s <- as.data.frame(t(abu_dino))
index <- as.data.frame(colnames(s))
index$phylum <- mapfile[index$`colnames(s)`, 'Genus']
colnames(s) <- 1:length(s)

#melt for bar plot

s.melt <- melt(s)
s.melt$SampleID <- rownames(s)
s.melt <- s.melt %>% filter(value != 0)
s.melt$taxa <- index[s.melt$variable, 'phylum']
subjmeans <- cast(s.melt, SampleID~taxa, sum)

write.csv(subjmeans, file="18S_dino_genus.csv")

#make version filled so all out of 1 (total dino proportion)
subjmeans_prop <- data.stand((as.matrix(subjmeans)), method='total', margin='row', plot=F)
write.csv(subjmeans_prop, file="18S_dino_genus_proportional.csv")



```
top 20 taxa heatmap

```{r}
r <- unique.select
colnames(r) <- mapfile[colnames(r), 'Taxa']
   
# 
r.fix <-as.data.frame(t(rowsum(t(r), group = colnames(r), na.rm = T)))
# r.fix <- sapply(split.default(r, names(r)), rowSums, na.rm = TRUE) %>%as.data.frame()
# r.2<- sort(colSums(r.fix), decreasing =TRUE)


#try heatmaps with melt/cast to get combined taxa names (no repeats)
# s <- unique.select
# index <- as.data.frame(colnames(s))
# index$taxa <- mapfile[index$`colnames(s)`, 'Taxa']
# colnames(s) <- 1:length(s)
# 
# s.melt <- melt(s)
# s.melt$SampleID <- rownames(s)
# s.melt <- s.melt %>% filter(value != 0)
# s.melt$taxa <- index[s.melt$variable, 'taxa']
# 
# s.melt <- s.melt %>%
#    group_by(SampleID, taxa) %>%
#   mutate(total_mmol = sum(value))
# 
# subjmeans <- cast(s.melt, SampleID~taxa, sum)
# 
# r<-subjmeans
# row.names(r) <- subjmeans$SampleID
# r <- r %>%
#    select(-SampleID)

#save for correlation top 99.9%
collapsed.select <- r.fix[,colMeans(r.fix) > 0.001]
 write.csv(collapsed.select, file="18S_taxa_correlation.csv")

#save for CCA top taxa
 metab_cca <- r.fix[, colSums(metab_cca!=0) > 0]
  top_20 <- names(sort(colSums(r.fix), decreasing =TRUE)[1:20])
  tally.top <- r.fix[top_20] 
  
  write.csv(tally.top, file="18S_taxa_top20.csv")


hm <- function(x,y) {
x <- x[, colSums(x!=0) > 0]
top_20 <- names(sort(colSums(x), decreasing =TRUE)[1:20])
tally.top <- x[top_20]
pheatmap(t(tally.top), main = y, cex = 1)
}

hm(r.fix, "18S Most Abundant Taxa")


#all with bray clustering
r <- unique.select 
colnames(r) <- mapfile[colnames(r), 'Taxa']

hm <- function(x,y) {
x <- x[, colSums(x!=0) > 0]
top_20 <- names(sort(colSums(x), decreasing =TRUE)[1:20])
tally.top <- x[top_20]
siteDist <- vegdist(t(tally.top), "bray", diag = TRUE)
siteDistcol<-vegdist(tally.top, 'bray', diag=TRUE)
pheatmap(t(tally.top), main = y, cex = 1, clustering_distance_rows = siteDist, clustering_distance_cols = siteDistcol)
}

EukTop20 <- hm(r.fix, "18S Most Abundant Taxa")

save_pheatmap_pdf <- function(x, filename, width=10, height=10) {
  stopifnot(!missing(x))
  stopifnot(!missing(filename))
  pdf(filename, width=width, height=height)
  grid::grid.newpage()
  grid::grid.draw(x$gtable)
  dev.off()
}
save_pheatmap_pdf(EukTop20, "EukTop20_all_bray.pdf")

#just incubation
 r.inc <- r[grep('ppt', rownames(r)),]

hm <- function(x,y) {
x <- x[, colSums(x!=0) > 0]
top_20 <- names(sort(colSums(x), decreasing =TRUE)[1:20])
tally.top <- x[top_20]
pheatmap(t(tally.top), main = y, cex = 1)
}

hm(r.inc, "18S Most Abundant Taxa")

#just field with bray clustering
  r.field <- r.fix[grep('B|Ev|Hero', rownames(r.fix)),] 

hm <- function(x,y) {
x <- x[, colSums(x!=0) > 0]
top_20 <- names(sort(colSums(x), decreasing =TRUE)[1:20])
tally.top <- x[top_20]
siteDist <- vegdist(t(tally.top), "bray", diag = TRUE)
siteDistcol<-vegdist(tally.top, 'bray', diag=TRUE)
pheatmap(t(tally.top), main = y, cex = 1, clustering_distance_rows = siteDist, clustering_distance_cols = siteDistcol)
}

EukTop20 <- hm(r.field, "18S Most Abundant Taxa")

save_pheatmap_pdf <- function(x, filename, width=10, height=10) {
  stopifnot(!missing(x))
  stopifnot(!missing(filename))
  pdf(filename, width=width, height=height)
  grid::grid.newpage()
  grid::grid.draw(x$gtable)
  dev.off()
}
save_pheatmap_pdf(EukTop20, "EukTop20_field_bray_fix.pdf")


#just inc with bray clustering
 r.inc <- r.fix[grep('ppt', rownames(r.fix)),]
 
hm <- function(x,y) {
x <- x[, colSums(x!=0) > 0]
top_20 <- names(sort(colSums(x), decreasing =TRUE)[1:20])
tally.top <- x[top_20]
siteDist <- vegdist(t(tally.top), "bray", diag = TRUE)
siteDistcol<-vegdist(tally.top, 'bray', diag=TRUE)
pheatmap(t(tally.top), main = y, cex = 1, clustering_distance_rows = siteDist, clustering_distance_cols = siteDistcol)
}

EukTop20 <-hm(r.inc, "18S Most Abundant Taxa")


save_pheatmap_pdf <- function(x, filename, width=10, height=10) {
  stopifnot(!missing(x))
  stopifnot(!missing(filename))
  pdf(filename, width=width, height=height)
  grid::grid.newpage()
  grid::grid.draw(x$gtable)
  dev.off()
}
save_pheatmap_pdf(EukTop20, "EukTop20_inc_bray_fix.pdf")
```

diversity indexes with anova

```{r}
library(vegan)
library(ggplot2)

#restrict metadata to useful samples 

data <- metadata[grep('Station|ppt|Core|Hero', rownames(metadata)),]
data[,1:3] <- NULL
#data <- data[-1,] 
# data <- data[-31,] 

#shannon diversity index
shann <- diversity(unique.s) #use data from BEFORE normalizing

# Simpson index
simp <- diversity(unique.s, "simpson")  

#inverse Simpson
insimp <- diversity(unique.s, "invsimpson")

par(mfrow = c(1, 3))  # to generate panels with 1 row of 3 graphs
hist(shann)
hist(simp)
hist(insimp)

#add inverse simpson diversity metric (looked best) to metadata
a <- as.data.frame(insimp)
data$InvSimpson  <- a[rownames(data), 'insimp']

#write out csv of wide metadata+diversity for CCA
write.csv(data, file="18S_meta.csv")

#plots divided by different sampling metrics (not sure which you think is most important)
ggplot(data, aes(x = as.factor(TempC), y = InvSimpson, fill=Sample_type)) + 
  geom_boxplot()

ggplot(data, aes(x = as.factor(Salinity), y = InvSimpson, fill=Sample_type)) + geom_boxplot()


invSimp<-ggplot(data, aes(x = as.factor(sample_type), y = InvSimpson, fill=sample_type)) + geom_boxplot()
invSimp
save_plot("InvSimp_18S.pdf", invSimp, base_height = 10, base_width = 10, units="in")

anova_result <- aov(InvSimpson ~ as.factor(TempC), data)
summary(anova_result)

anova_result_sal <- aov(InvSimpson ~ as.factor(Salinity), data)
summary(anova_result_sal)

anova_result_samp <- aov(InvSimpson ~ as.factor(sample_type), data)
summary(anova_result_samp)
capture.output(summary(anova_result_samp), file="ANOVA_18SSimpson.csv")

#Post-hoc Tukey's HSD if ANOVA interaction significant
TukeyHSD(anova_result_samp)
TKHSD_Compound <- TukeyHSD(anova_result_samp)
capture.output(TKHSD_Compound, file = "TKHSD_18SSimpson.csv")



#try with inc samples only
data_inc <- data %>%
  filter(str_detect(Better_SampID, "ppt"))
anova_result_samp <- aov(InvSimpson ~ as.factor(sample_type), data_inc)
summary(anova_result_samp)


#try with field samples only
data_field <- data %>%
  filter(!str_detect(Better_SampID, "ppt"))
anova_result_samp <- aov(InvSimpson ~ as.factor(sample_type), data_field)
summary(anova_result_samp)

#not significant (barely), could run a Tukey (the negative temp looks like it could be significant)

```

CA

```{r}
library(ca)
library(factoextra)
#hellinger transform data
Hellinger_unique <- decostand(unique.s, method = "hellinger", na.rm = TRUE) 
Hellinger_unique <- Hellinger_unique[ order(row.names(Hellinger_unique)), ]
#ca
unique.select.ca <- ca(Hellinger_unique)
unique.select.ca.var <- get_eigenvalue(unique.select.ca)
unique.select.ca.res <- get_ca_col(unique.select.ca)
samples.x <- unique.select.ca$rowcoord[,1]
samples.y <- unique.select.ca$rowcoord[,2]

dim.1.var <- round(unique.select.ca.var$variance.percent[1], 1)
dim.2.var <- round(unique.select.ca.var$variance.percent[2], 2)
envi_color<-c(seawater='purple', low='orange', mid='green', high='pink')
plot(samples.x, samples.y,
ylab = paste0('Dim 2: ', dim.2.var, '%'),
xlab = paste0('Dim 1: ', dim.1.var, '%'),
pch = 21,col="black", cex=1, bg=envi_color[data$location])
row$coord

#remove outliers


```






NMDS 

```{r}
library(phyloseq)
####################### Ordination Plots ############################
Hellinger_unique <- decostand(unique.s, method = "hellinger", na.rm = TRUE) 

#transform dataframes into matrices, switching the rows and columns for downstream analyses
Hellinger_unique <- t(Hellinger_unique) #flip

#create OTU matrix
OTUmat <- as.matrix(Hellinger_unique)
#create taxa matrix
taxmat <- as.matrix(mapfile[,4])

#combine into a phyloseq object
OTU = otu_table(OTUmat, taxa_are_rows = TRUE)
TAX = taxa_names(taxmat)
sampledata <- data
physeq = phyloseq(OTU,TAX,sampledata)
physeq

#ordination 
theme_set(theme_bw())
ord2 <- ordinate(physeq, "NMDS", "bray")

plot_ordination(physeq, ord2, type="sample_type", color = "TempC") +       geom_point(size=3) + 
   facet_wrap(~TempC, labeller = depth_labeller) +            theme(legend.position = "top", legend.title = element_blank()) + scale_shape_discrete(labels = c("Shaved", "Not Shaved"))


#permutation MANOVA vegan package
df = data.frame(sample_data(physeq))
d = phyloseq::distance(physeq, "bray")

adonis(formula = d ~ treatment*shave + depth, data = df) 
#Homogeneity of dispersion test
beta <- betadisper(d, df$treatment)
permutest(beta) #betadisper results are not sig, meaning we reject null hypothesis that tour groups have the same dispersions... confident in result

#p.adj to correct for multiple comparisons
p_perm <- c(0.001, 0.0152, 0.001, 0.133)
p.adjust(p_perm, "bonferroni")


```
